---
title: "Tips-data [A3]"
---

# Tips data

```{r}
#| label: setup

library(tidyverse)
library(mosaic)
library(ggformula)
library(crosstable)
library(broom)
```

```{r}
mydata <- read_csv("../../data/tip.csv")
mydata


```

```{r}
mydata %>% 
  crosstable(Tip ~ Preferance) %>% 
  as_flextable()
```

```{r}
tip_modified <- mydata %>% 
  dplyr::mutate(
    Gender = as_factor(Gender),
    Preferance = as_factor(Preferance)
  )

glimpse(tip_modified)
```

```{r}
shapiro.test(tip_modified$Tip)

```

```{r}
mosaic::t_test(Tip ~ Preferance, data = tip_modified) 

```

```{r}
wilcox.test(Tip ~ Preferance, data = tip_modified) %>%  
  broom::tidy()
```

Since the **p-value is high**, we shall not reject the null hypothesis. Any observed difference might be due to chance, and there is no significant difference between the groups - Non-veg and Veg.

```{r}
tips_permute <-
  do(4999) * diffmean(data=tip_modified, Tip~shuffle(Preferance))
tips_permute
```

```{r}
tip_modified %>%
  gf_histogram(~Tip, fill = ~Preferance) 
```

```{r}
tip_modified %>%
  gf_density(~Tip,
             fill = ~Gender,
             alpha = 0.8,
             title = "DENSITY PLOT"
    
  )
```

-   This data is **not normally distributed**. A normal distribution would depict a bell curve. Here, the tips are heavily concentrated in the lower values, and there’s a long right tail. Hence, the data is heavily skewed to the right for both genders.

-   Overall, tips are typically low, with a few outliers where people gave much higher amounts, especially among males.
