---
title: "Day-4"
author: "Aish"
---

# Day 4

## Quantities

We have downloaded a new library today - "crosstable". It is used for fast stats for multiple variables in table form.

```{r}
#| label: setup
library(tidyverse)
library(mosaic)
library(ggformula)
library(skimr)
library(crosstable)
```

## Examining the 'Diamonds' dataset.

```{r}
glimpse(diamonds)
```

```{r}
skim(diamonds)
```

```{r}
inspect(diamonds)
```

```{r}
gf_histogram(~price, data = diamonds) %>%
  gf_labs(
    title = "Plot 1A: Diamond Prices",
    caption = "ggformula"
  )
```

```{r}
diamonds %>% 
  gf_histogram(~price, bins = 100) %>% 
  gf_labs(
    title = "Plot 1B: Diamond Prices",
    caption = "ggformula"

  )
```

#### Insights:

-   There are a great many diamonds at relatively low prices, but there are a good few diamonds at very high prices too.

-   Using a high number of bins does not materially change the view of the histogram.

```{r}
diamonds %>% 
  gf_histogram(~carat) %>% 
  gf_labs(
     title = "Plot 2A: Carats of Diamonds",
    caption = "ggformula"
  )
```

```{r}
diamonds %>% 
  gf_histogram(~carat, bins = 100) %>% 
  gf_labs(
     title = "Plot 2B: Carats of Diamonds",
    caption = "ggformula"
  )
```

#### insights:

-   there is a marked “discreteness” to the distribution. Some values of carat are far more common than others. For example, 1, 1.5, and 2 carat diamonds are large in number.

```{r}
diamonds %>% 
  gf_histogram(~price, fill = ~cut) %>% 
  gf_labs(
     title = "Plot 3A: Diamond Prices", caption = "ggformula")
```

```{r}
diamonds %>%
  gf_histogram(~price, fill = ~cut, color = "black", alpha = 0.3) %>%
  gf_labs(
    title = "Plot 3B: Prices by Cut",
    caption = "ggformula"
  )
```

```{r}
diamonds %>%
  gf_histogram(~price, fill = ~cut, color = "black", alpha = 0.3) %>%
  gf_facet_wrap(~cut) %>%
  gf_labs(
    title = "Plot 3C: Prices by Filled and Facetted by Cut",
    caption = "ggformula"
  ) %>%
  gf_theme(theme(
    axis.text.x = element_text(
      angle = 45,
      hjust = 1
    )
  ))
```

```{r}
diamonds %>%
  gf_histogram(~price, fill = ~cut, color = "black", alpha = 0.3) %>%
  gf_facet_wrap(~cut, scales = "free_y", nrow = 2) %>%
  gf_labs(
    title = "Plot 3D: Prices Filled and Facetted by Cut",
    subtitle = "Free y-scale",
    caption = "ggformula"
  ) %>%
  gf_theme(theme(
    axis.text.x =
      element_text(
        angle = 45,
        hjust = 1
      )
  ))
```

```{r}
#| eval: false
install.packages("shiny")
library(shiny)
runExample("01_hello") # an interactive histogram
```

## Examining the 'Race' dataset.

```{r}
race_df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-26/race.csv")
rank_df <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-10-26/ultra_rankings.csv")
```

```{r}
glimpse(race_df)
```

```{r}
glimpse(rank_df)
```

```{r}
race_df %>%
  favstats(~distance, data = .)
```

```{r}
##
race_df %>%
  favstats(~participants, data = .)
```

```{r}
##
rank_df %>%
  drop_na() %>%
  favstats(time_in_seconds ~ gender, data = .)
```

```{r}
## library(crosstable)
crosstable(time_in_seconds + age ~ gender, data = rank_df) %>%
  crosstable::as_flextable()
```

### **Max. Races and participants**

```{r}
race_df %>%
  count(country) %>%
  arrange(desc(n))
```

```{r}
rank_df %>%
  count(nationality) %>%
  arrange(desc(n))
```

Insights:

-   the first piece of code tells the no. of races hosted by the respective countries, whereas the secodn piece of code tells no, of participants from each county.

-   The top three locations for races were the USA, UK, and France. These are also the countries that send the maximum number of participants, naturally!

### **Max. Winners**

```{r}
rank_df %>%
  filter(rank %in% c(1, 2, 3)) %>%
  count(nationality) %>%
  arrange(desc(n))
```

### **Which countries have had the most top-3 finishes?**

```{r}
longest_races <- race_df %>%
  slice_max(n = 5, order_by = distance) # Longest distance races
longest_races
longest_races %>%
  left_join(., rank_df, by = "race_year_id") %>% # total participants in longest 4 races
  filter(rank %in% c(1:10)) %>% # Top 10 ranks
  count(nationality) %>%
  arrange(desc(n))
```

### **What is the distribution of the finishing times?**

```{r}
rank_df %>%
  gf_histogram(~time_in_seconds, bins = 75) %>%
  gf_labs(title = "Histogram of Race Times")
```

-   "1e+05" means 1×10 to the power 5. this is equal to 100,000 seconds.

### **What is the distribution of race distances?**

```{r}
race_df %>%
  gf_histogram(~distance, bins = 50) %>%
  gf_labs(title = "Histogram of Race Distances")
```

```{r}
race_df %>%
  filter(distance == 0)
```

-   some of these zero-distance races have had participants too! Perhaps these were cancelled events…all of them are stated to be `100 mile` events…

### **What is the distribution of finishing times for race distance around 150?**

```{r}
race_times <- race_df %>%
  count(start_time) %>%
  arrange(desc(n))
race_times
```

```{r}
# Demo purposes only!

race_start_factor <- race_df %>%
  filter(distance == 0) %>% # Races that actually took place
  mutate(
    start_day_time =
      case_when(
        start_time > hms("02:00:00") &
          start_time <= hms("06:00:00") ~ "early_morning",
        start_time > hms("06:00:01") &
          start_time <= hms("10:00:00") ~ "late_morning",
        start_time > hms("10:00:01") &
          start_time <= hms("14:00:00") ~ "mid_day",
        start_time > hms("14:00:01") &
          start_time <= hms("18:00:00") ~ "afternoon",
        start_time > hms("18:00:01") &
          start_time <= hms("22:00:00") ~ "evening",
        start_time > hms("22:00:01") &
          start_time <= hms("23:59:59") ~ "night",
        start_time >= hms("00:00:00") &
          start_time <= hms("02:00:00") ~ "postmidnight",
        .default = "other"
      )
  ) %>%
  mutate(
    start_day_time =
      as_factor(start_day_time) %>%
        fct_collapse(
          .f = .,
          night = c("night", "postmidnight")
        )
  )
##
# Join with rank_df
race_start_factor %>%
  left_join(rank_df, by = "race_year_id") %>%
  drop_na(time_in_seconds) %>%
  gf_histogram(
    ~time_in_seconds,
    bins = 75,
    fill = ~start_day_time,
    color = ~start_day_time,
    alpha = 0.5
  ) %>%
  gf_facet_wrap(vars(start_day_time), ncol = 2, scales = "free_y") %>%
  gf_labs(title = "Race Times by Start-Time")
```

## **Distributions and Densities in the Wild**

```{r}
pop <- read_csv("../../data/populations.csv")
pop
inspect(pop)
```

```{r}
glimpse(pop)
```

```{r}
##
gf_histogram(~value, data = pop, title = "Long Tailed Histogram")
##
gf_density(~value, data = pop, title = "Long Tailed Density")
```

#### insights

-   I don't really understand much from these graphs.

```{r}
gf_histogram(~ log10(value), data = pop, title = "Histogram with Log transformed x-variable")
##
gf_density(~ log10(value), data = pop, title = "Density with Log transformed x-variable")
```
